---
title: "Simple Search Writeup"
author: "Dalton Gusaas && Thomas Harren"
output: 
  html_document:
    toc: true
---

# Introduction

In these set of experiements we explore the differences between three different techniques. Full explantions can be found on our readme at https://github.com/harre096/simple-search/blob/master/docs/Tweak.md. 

As a summary, our techiques are: 

* half_random_half_hill_climber
    * Runs random search for half of the tries and then uses hill_climber using the flip-one-bit mutator. 
* rand_reset_10_times
    * Runs the above 10 times, generating a new seed each time.
* random_search
    * Nic's random search method.

The mutator that we use is flip-one-bit, which modifies an answer by randomly flipping one bit.

We apply these techniques to seven data sets: 

* "knapPI_11_20_1000_4" 
* "knapPI_13_20_1000_4" 
* "knapPI_16_20_1000_4"
* "knapPI_11_200_1000_4" 
* "knapPI_13_200_1000_4"
* "knapPI_16_200_1000_4"


There are two possible variables to the trials:
* Varying the number of repetitions
    * We set this as 30 each time. We thought that this seemed like a fairly large number of repetitions.
* Varying the number of tries that each run gets to find the answer. (These tries were split between random and our hillclimb in the various ways described above.) W tested with 3 different tries amounts:
    * 10,000
    * 100,000
    * 300,000
    
#Let's look at some data!

## 10,000 tries
```{r warning=FALSE}
library("ggplot2")

data_50_runs <- read.csv("10000Tries.txt", sep="")

ggplot(data_50_runs, 
       aes(x=factor(Max_evals), y=Score, group=Max_evals)) + 
  geom_boxplot() + facet_grid(Search_method ~ Problem)
```

```{r}
library("rpart")
library("rpart.plot")

rp <- rpart(Score ~ Search_method + Problem + Max_evals, data=data_50_runs)

rpart.plot(rp, type=3, extra=100)
```

## 100,000 tries
```{r warning=FALSE}
library("ggplot2")

data_50_runs <- read.csv("100000Tries.txt", sep="")

ggplot(data_50_runs, 
       aes(x=factor(Max_evals), y=Score, group=Max_evals)) + 
  geom_boxplot() + facet_grid(Search_method ~ Problem)
```

```{r}
library("rpart")
library("rpart.plot")

rp <- rpart(Score ~ Search_method + Problem + Max_evals, data=data_50_runs)

rpart.plot(rp, type=3, extra=100)
```

## 300,000 tries
```{r warning=FALSE}
library("ggplot2")

data_50_runs <- read.csv("300000Tries.txt", sep="")

ggplot(data_50_runs, 
       aes(x=factor(Max_evals), y=Score, group=Max_evals)) + 
  geom_boxplot() + facet_grid(Search_method ~ Problem)
```

```{r}
library("rpart")
library("rpart.plot")

rp <- rpart(Score ~ Search_method + Problem + Max_evals, data=data_50_runs)

rpart.plot(rp, type=3, extra=100)
```

# Conclusion

From these results its clear that ...